Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram

### 2015-11-20 papers/MbhMsph/mm.tex
victory2

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$  
and define a tight 
\emph{red sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$, 
i.e.~a $\pm 2\sigma$ range of 0.1--5\%.
At the low-mass end, the bulges of late-type (Sp) galaxies define a much steeper 
\emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by the current bulge sample with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies 
(most elliptical galaxies contain disks, i.e.~$L_{\rm sph} \neq L_{\rm gal}$) --
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

{\bf 
As a final remark, we comment on the work done by \cite{reinesvolonteri2015}, 
who investigated the relationship between black hole mass and total galaxy stellar mass, $M_{\rm *,gal}$. 
Their Figure 8 presents the $M_{\rm BH} - M_{\rm *,gal}$ distribution 
for a sample of $\approx 260$ local AGNs with virial black hole masses  
and for $\approx 80$ galaxies with dynamical black hole masses. 
They concluded that the AGN sample and the early-type galaxies with quiescent black holes define two distinct sequences 
in their $M_{\rm BH} - M_{\rm *,gal}$ diagram; 
these two sequences have similar slope, but a normalization factor different by more than one order of magnitude.  
Since we noted that the \cite{jiang2011a} AGN sample follows the steeper $M_{\rm BH} - M_{\rm *,sph}$ correlation 
traced by our spiral galaxy bulges (the majority of which host an AGN), 
it would be interesting to recover the $M_{\rm BH} - M_{\rm *,sph}$ distribution 
also for the AGN sample of \cite{reinesvolonteri2015}. \\
However we do note that there is emerging evidence (e.g.~\citealt{busch2015,subramanian2015}) 
for a population of bulges with black hole masses residing below (or to the right of) the red and blue $M_{\rm BH} - M_{\rm *,sph}$ 
sequences constructed here using samples with directly measured black hole masses, as speculated by Batcheldor (2010).
}

@@

      Although we do not consider the S\'ersic index a good indicator as of the nature of a bulge (e.g.~\citealt{grahamworley2008}), 

@@

      spheroids with $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 

>>>>

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, $M_{\rm BH}$, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between $M_{\rm BH}$ 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
While most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies,
our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$  
and define a tight 
\emph{red sequence} with intrinsic scatter $\epsilon_{(M_{\rm BH}|M_{\rm *,sph})} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$, 
i.e.~a $\pm 2\sigma$ range of 0.1--5\%.
At the low-mass end, the bulges of late-type galaxies define a much steeper 
\emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$ and $M_{\rm BH}/M_{\rm *,sph}$ equal to $0.02\%$ at $M_{\rm BH} \approx 10^6~\rm M_\odot$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) our S\'ersic galaxy sample follows $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, 
a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by the current bulge sample with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies 
-- once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

{\bf 
As a final remark, we comment on the work by \cite{reinesvolonteri2015}, 
who investigated the relationship between black hole mass and total galaxy stellar mass, $M_{\rm *,gal}$. 
Their Figure 8 presents the $M_{\rm BH} - M_{\rm *,gal}$ distribution 
for a sample of $\approx 260$ local AGNs with virial black hole masses  
and for $\approx 80$ galaxies with dynamical black hole masses. 
They concluded that the AGN sample and the early-type galaxies with quiescent black holes define two distinct sequences 
in their $M_{\rm BH} - M_{\rm *,gal}$ diagram; 
these two sequences have similar slope, but a normalization factor different by more than one order of magnitude.  
Since we noted that the \cite{jiang2011a} AGN sample follows the steeper $M_{\rm BH} - M_{\rm *,sph}$ correlation 
traced by our spiral galaxy bulges (the majority of which host an AGN), 
it would be interesting to recover the $M_{\rm BH} - M_{\rm *,sph}$ distribution 
also for the AGN sample of \cite{reinesvolonteri2015}. \\
However we do note that there is emerging evidence (e.g.~\citealt{busch2015,subramanian2015}) 
for a population of bulges with black hole masses residing below (or to the right of) the red and blue $M_{\rm BH} - M_{\rm *,sph}$ 
sequences constructed here using samples with directly measured black hole masses, as speculated by Batcheldor (2010).
}

@@

      Although we do not consider the S\'ersic index a good indicator of the nature of a bulge (e.g.~\citealt{grahamworley2008}), 

@@

      spheroids with $n_{\rm sph}<2$ 
